Point transect distance sampling in inlabru

### 2019-09-02 paper.tex
edit case study

% RJC 09/08/2019 Added case study text between % addtextstart and % addtextend markers.
% addtextstart
In this section, we present a description of Hawaiian forest birds, forest bird monitoring, the study species, study area, bird sampling, and description of the data. \hawaii{} is home to the largest adaptive radiation of oceanic archipelago birds worldwide. The Hawaiian Honeycreepers (Drepanidinae) is comprised of 21 genera and more than 50 species generally grouped according to feeding niches with many species and subspecies being island endemics. The honeycreepers, like most insular bird species, evolved life history strategies that include greater longevity with high adult survival, delayed maturation with an extended post-fledging dependency period, decreased clutch size and fecundity, and reduced dispersal abilities with high site fidelity \citep{woodworth_demography_2009}. While the honeycreepers diversified into the plethora of ecological niches they also had a loss or reduction of defences against competitors, predators and diseases. The loss of defensive adaptations made the honeycreepers particularly vulnerable to humans and the changes and threats they brought. As a result, more than half of the honeycreepers are extinct \citep{banko_extinctions_2009}. A mere 23 honeycreeper species persist with most species restricted to minute portions of their original ranges and in small numbers. The IUCN lists 21 honeycreeper species as critically endangered, endangered or vulnerable, and only two species are categorized as least concern \citep{iucn_redlist_2018}.

In the late $19^{\textnormal{th}}$ century naturalists to \hawaii{} described the Hawaiian forest birds and documented species distributions and abundances. Subsequent naturalists to the islands documented range contractions and population declines of native species while noting expansion in non-native bird species ranges and abundances. It was not until standardized and statistically robust sampling techniques to estimate population sizes and associated standard errors were initiated that the extent to which honeycreepers had declined was appreciated. Large-scale, quantitative surveys of Hawaiian forest birds and their habitat commenced in the mid-1970s through the Hawaii Forest Bird Survey (HFBS) \citep{scott_HFBS_1986}. Information from the HFBS was used to update the listing and delisting of endangered species, and establish preserves that coincided with native bird hotspots, including Hakalau Forest National Wildlife Refuge that was the first wildlife refuge with the primary purpose to protect, conserve and manage native forests for threatened and endangered bird and plant species. Due to the broad-scale coverage and robust design HFBS has become the baseline to determine changes in bird species distributions, population sizes and trends in density patterns over time.

\subsection*{Study species}
The \hawaii{} \akepa{} (hereafter \akepa{}; \textit{Loxops coccineus}; nomenclature according to \citealp{usfws_akepa_1970}) is an internationally and federally endangered Hawaiian honeycreeper (\citealp{usfws_akepa_1970, birdlife_akepa_2016}) that is endemic to \hawaii{} Island, USA. \Akepa{} is a foraging specialist whose diet consists of spiders, insects and psyllids, as well as nectar. \Akepa{} uses its cross-bill to probe and open terminal leaf clusters and extracts prey with its tongue. During the 20th century \akepa{} declined dramatically due to habitat modification \citep{scott_HFBS_1986, pratt_avifaunal_1994},  mosquito-transmitted avian diseases \citep{pratt_avifaunal_1994, atkinson_wildlife_1995}, introduced predators \citep{lepson_akepa_1997}, and food resources competitors \citep{lepson_akepa_1997}. \Akepa{} has a global abundance of 16,248 (95\%CI 10,074\textendash25,198) birds that has been restricted to five spatially distinct populations \citep{judge_akepa_2018}. Hakalau Forest National Wildlife Refuge on \hawaii{} Island (hereafter Hakalau; 19$^{\circ}$ 51\textquotesingle N, 155$^{\circ}$ 18\textquotesingle W), supports the largest \akepa{} population that in 2012 was estimated at more than 11,000 birds \citep{camp_statespace_2016}. Maintaining and expanding the \akepa{} population at Hakalau is a primary conservation concern. Moreover, having unbiased and precise abundance estimates are required by land and resource managers for evaluating management actions and establishing management planning, and policy makers for decision-making processes.

\subsection*{Study area}
Hakalau was established in 1985 to conserve 15,390-ha of montane forest habitat for native forest birds and rainforest plants. Annual forest bird surveys were initiated in 1987 to determine population status and track trends in abundance. Survey points were established along 15 transects following a systematic, random design with points approximately 150-m apart on transects located either 500 or 1,000 m apart. Following \cite{camp_population_2010, camp_statespace_2016}, we restrict our study area to the 3,061 ha open-forest stratum of Hakalau. The study area is between 1,400 and 1,920 m elevation of montane forest dominated by native \textquotesingle \={o}hi\textquotesingle a (\textit{Metrosideros polymorpha}) and koa (\textit{Acacia koa}) with a mixture of native and non-native understory plants. The open-forest stratum was previously heavily grazed, and since the removal of cattle in 1988 regeneration has proceeded naturally \citep{maxfield_hakalau_1998}. To the north the study area follows the refuge boundary while to the east it is bounded by a fence line (Fig. \ref{fig:2002studyareapointspt}). The southern boundary was modified from \cite{camp_population_2010} to exclude the non-sampled forested portion of their study area. The west side of the study area is bounded by pasture that is dominated by grass and is unsuitable habitat for \akepa. The substrate in the southern portion of the study area is nutrient-rich, old lava flows (11,000 to 64,000 yrs old) whereas the northern portion occurs on new lava flows (5,000 to 10,000 yrs old) that are nutrient depleted \citep{ccp_hakalau_2010}.

@@

\subsection*{Bird sampling}
Surveys used point-transect methods, a form of distance sampling, where horizontal distances from survey points to individually detected birds are used to model a species-specific detection probability that is used to estimate absolute abundance and variance \citep{buckland_distance_2015}. Surveys commenced at dawn and continued until 11:00 or halted when weather conditions exceeded prescribed conditions that hindered detecting birds (light rain, and wind and gust strength \textgreater Baufort scale 3). During 8-min counts trained observers recorded the species, exact distance to the nearest meter and detection type for each bird detected, along with the sampling conditions cloud cover, rain, wind strength, gust strength, and time of day each point was surveyed.

\subsection*{Data description}
For purposes of our analysis, we selected a single survey from the \akepa{} time series based on broad sampling of the study area and sufficient numbers of detections. In 2002, 195 points were sampled using point-transect distance sampling methods within the 3,061 ha open-forest study area of Hakalau (Fig. \ref{fig:2002studyareapointspt}). On 67 points 152 \akepa{} were detected. The number of detections ranged from zero to 6 (Fig. \ref{fig:2002countspt}). These data are used for both the two-stage and one-stage model-based analyses. \cite{camp_population_2010,camp_statespace_2016} provide a detailed description of Hakalau, the open-forest study area and the bird surveys.

\begin{figure}
	\centering
	\includegraphics[scale=0.5]{figures/2002counts_pt}
	\caption{Counts of \akepa{} by point during the 2002 survey.}
	\label{fig:2002countspt}
\end{figure}

@@

In this section we introduce \texttt{inlabru} syntax and provide example code to demonstrate to specify the model.  First we specify the SPDE model as
\begin{lstlisting}
matern <- inla.spde2.pcmatern(mesh, 
                              prior.sigma = c(2, 0.01),    
                              prior.range = c(130, 0.01))
\end{lstlisting}
which uses the penalised complexity priors for the hyperparameters of the GMRF as introduced by \cite{simpson_penalising_2017}, setting $\mathbb{P}(\sigma > 2) = 0.01$ and $\mathbb{P}(\rho < 130) = 0.01$ where $\rho$ is the range of the Mat\'ern covariance.

The data consists only of distances from the observer so we use equation \eqref{intensity-incomplete} to specify the log thinning probability function as
\begin{lstlisting}
thinning_function <- function(distance, lsig){
    log(2*pi) + log(distance) + log_halfnormal(distance, lsig)
    }
\end{lstlisting}
where \texttt{log\_halfnormal} is the log half-normal detection function $\log g(r | \sigma) = -1/2(r/\sigma)^2$ and the argument \texttt{lsig} is $\log(\sigma)$.  The column in the dataset containing the observed distances is named \texttt{distance} amd the software will search for the appropriate data using the names of the arguments specified in this function.  To allow the software to distinguish between names of covariates and parameters to be estimated we must also specify a model components object
\begin{lstlisting}
cmp <- ~ Intercept + grf(map = coordinates, model = matern) + lsig 
\end{lstlisting} 
This object names all model components to be estimated.  Note that the \texttt{grf} component is a name we have chosen and not a function.  The software allows user-specified names for the \texttt{f()} function known to users of \texttt{R-INLA}. \texttt{coordinates} is a keyword that allows users to supply their data in the form of a spatial object using the \texttt{sp} package \citep{pebesma_spatial_2005}.

The model formula object is
\begin{lstlisting}
fml <- coordinates + distance ~ Intercept + grf +
	thinning_function(distance, lsig)
\end{lstlisting}
To fit the model we specify starting values for the parameters, the domain of possible observed distances and fit the model using the \texttt{lgcp} function
\begin{lstlisting}
starting_values <- data.frame(grf = 0, lsig = 3.65, Intercept = 0)
W <- 60   # transect radius
distance_domain <- seq(.Machine$double.eps, W, length.out = 30)
lgcp(components = cmp, 
     data = obs,
     samplers = samplers,
     domain = list(distance = distance_domain),
     formula = fml,
     options = list(result = starting_values, max.iter = 30))
\end{lstlisting}
where our observed data is contained in a \texttt{SpatialPointsDataFrame} object named \texttt{obs}.  Since not all sampling units observed birds we also specify the \texttt{samplers} object, also a \texttt{SpatialPointsDataFrame}, which contains the locations of all sample units%.

@@

The predicted mean of the posterior intensity is shown in \autoref{fig:intensity-mean-cv} along with a map of the coefficient of variation (CV) of the posterior intensity field on a regular prediction grid with cell area of approximately 1.7 hectares.  This shows a region of high intensity in the south and much lower intensity in the north.  The CV plot shows clearly that there is less variation in the posterior intensity field in areas with greater sampling effort.  In general the south, with greater density of sampled locations, has a lower CV than the north, where sampling units are more spread out.  The areas of highest CV are those regions furthest away from any sampled location, reflecting the intuitive fact that we are least certain in areas further away from the sampled locations.

@@

The CV plot shows a measure of uncertainty relative to the mean intensity in each prediction grid cell.  However, often we are interested in the uncertainty around the mean intensity on the scale of the intensity function itself - perhaps seeking to identify areas where we can confidently predict particularly high or low intensities.  To attempt to communicate this uncertainty around the mean it is common [[references?]] to present the lower and upper quantiles corresponding to a 95\% credible interval for each prediction cell (\autoref{fig:intensity-quantiles}). [RJC - I don't think a reference is needed to substantiate presenting credible intervals or computing them using quantiles, particularly in a stats journal. AES - I was trying to find an example of someone presenting maps like these in a paper, yet to find one.  If you know of any please let me know!]

@@

This difference becomes clear if one generates realizations of the posterior intensity field.  Three such realizations are in \autoref{fig:intensity-realizations}.  There are some notable differences when compared to the mean and quantile summary maps.  First, each realization has finer-grained spatial structure than the mean of the field.  In practice, this indicates that we should expect more spatial structure in our collected data than what we might expect if we only look at the posterior mean of the field.  The mean field plot hides the spatial structure the model has been able to estimate from the data.  In general, the mean of a random field will contain less spatial structure than realisations of the field. [RJC - 'First' is not followed up with subsequent ordinal numbers; I suggest rewriting the above sentence to remove 'first' or add ordinal numbers below.]

@@

The quantile plots are more concerning.  As mentioned above, the temptation is to interpret these as possible intensities that produced our observed data.  However, they in fact show the quantiles of the \textit{marginal} probability density for each prediction location.  This essentially treats each prediction cell as independent of the others. Whilst any one prediction location may be well described by such quantiles, to present them side by side in one image obscures the fact that the probability that all (or some subset of) the prediction cells would \textit{simultaneously} achieve that value will be in fact much lower than the quantile value and in many cases would be essentially zero.  [RJC - I stumbled over this sentence several times and I think it needs to be rewritten.]  To demonstrate this we treated the lower and upper quantile plots as though they were intensities and integrated them to obtain an expected abundance estimate of approximately 3,300 for the 0.025 quantile and 12,400 for the 0.975 quantile.  A naive (and tempting) interpretation of these numbers is as upper and lower limits of the 95\% credible interval for abundance. However, these abundance estimates are essentially impossible according to the estimated posterior for abundance (see \autoref{fig:realized-abundance-posterior}).  This emphasises that the temptation to treat these maps as showing possible intensity surfaces leads to outcomes (and possible management decisions) that are inconsistent with the very model that generated the maps.

As an alternative, we suggest using excursion sets and excursion functions \citep{bolin_excursion_2015}.  This is a recently developed approach to representing uncertainty in stochastic processes that specifically considers joint probability of events across multiple locations.  The positive excursion set with level $u$ for a function $f(s)$ with domain $\Omega$ is $A_u^{+}(f) = \{ s \in \Omega ; f(s) > u \}$ i.e. the set of all locations in $\Omega$ where $f$ exceeds the a threshold value $u$. For a stochastic process $\lambda(s)$ the positive excursion set with level $u$ and probability $1 - \alpha$ is

\begin{equation*}
E_{u,\alpha}^{+}(\lambda) = \argmax_{D}\{\lvert D \rvert : \mathbb{P}\left[D \subset A_u^{+}(\lambda)\right] \geq 1 - \alpha \}
\end{equation*}
It is important to note that $A_u^{+}(f)$ specifies a set for which a function $f(s)$ exceeds a threshold value $u$ for \textit{every location} in the set.  This means that when considering random quantities such as $\lambda(s)$ the definition of the positive excursion set $E_{u,\alpha}^{+}(\lambda)$ is the largest such set for which a threshold is exceeded simultaneously for all locations in the set.  Because $\lambda(s)$ is stochastic we can only ever say this is true with some probability and use something akin to a confidence level by specifying $\alpha$.

Excursion sets can be estimated by considering candidate sets for $D$ of increasing size and a sequential integration scheme to estimate probabilities.  An implementation is available in the \texttt{excursions} package \citep{bolin_calculating_2018} available through the Comprehensive R Archive Network \citep{r_2017}.

\autoref{fig:excursions} (left panel) shows the positive excursion set with a level corresponding to 1 bird per hectare with probability 0.95.  This figure can be interpreted in a natural way.  It is the largest possible region for which the intensity is greater than 1 bird per hectare for every location within the region, with probability 0.95.

\begin{figure}
	\includegraphics[scale=0.5]{figures/excursions.png}
	\caption{Left:  The positive excursion set with a level corresponding to 1 bird per hectare and probability 0.95.  Right: The positive excursion function with a level corresponding to 1 bird per hectare}
	\label{fig:excursions}
\end{figure}

To visualise multiple such maps we can use the excursion function $F_u^{+}(s) = \sup \{1 - \alpha ; s \in E_{u,\alpha}^+ \}$, which defines for each location something similar to a p-value.  For each $s$ it is the largest possible probability $1 -\alpha$ for which that location would be in the excursion set defined using probability $1 - \alpha$.  The excursion function with a level corresponding to 1 bird per hectare is shown in \autoref{fig:excursions} (right panel).

@@

A key feature of these plots is that they display either discrete sub-regions of the area of interest (for excursion sets) or the probability of membership of such discrete sub-regions (for excursion functions).  We believe the risk of interpreting these maps as intensities is much lower than for the quantile plots.  The fact that it is possible to estimate discrete regions may also be useful to communicate results to conservation practitioners, land and resource managers and policy decision-makers.  We anticipate the excursion function will be most useful in communicating to audiences with statistical training. [RJC - I think statistically savvy audiences will readily grasp the excursion functions. Do you mean "without" statistical training?]

Another key aim of performing distance sampling surveys is to estimate the abundance of the population of interest.  By specifying our spatial model as a log-Gaussian Cox process, the posterior intensity is a random field rather than a deterministic function. Let $n$ denote the abundance of the population within the region of interest and $N$ its corresponding random variable within our Bayesian framework.  Integrating the mean of the random field provides a point estimate for the abundance, which we term the expected abundance.  Here we demonstrate the importance of considering the full posterior for $N$ which we call the \textit{realized} abundance.

Let $N = \int_{\Omega}\lambda(s)\mathrm{d}s$. Note that $N$ is a random variable and the integral of $\mathbb{E}[\lambda(s)]$ is equivalent to $\mathbb{E}(N)$ by Fubini's theorem, hence the term expected abundance.  As we mentioned above, this is the most sensible point estimate for the abundance.  We estimate the posterior $\pi(N | \lambda)$ by a monte-carlo method.  Taking $m$ monte-carlo samples  $\lambda_1, \ldots, \lambda_m$ of the posterior intensity field we estimate the posterior for the abundance as $\pi(N | \lambda) \approx 1 / m \sum_{i=1}^m \pi (N | \lambda = \lambda_i)$. This is well-defined since $\pi(N | \lambda = \lambda_i)$ is a Poisson probability mass function with rate parameter $\int_{\Omega}\lambda_i(s)\mathrm{d}s$. \autoref{fig:realized-abundance-posterior} shows the approximate posterior for $N$ with $m = 20,000$.  This allows us to estimate the probability for any specific value of \textit{realized} abundance $n$ by calculating $\pi(N = n | \lambda)$.  For reference, \autoref{fig:realized-abundance-comparison} shows a single Poisson PMF with rate parameter equal to the expected abundance.

@@

\begin{figure}
	\includegraphics[scale=0.6]{figures/halfnormal.png}
\end{figure}

@@

TODO HERE:  Something about how this is all a one-stage model and all these plots are based on sampling from the joint posterior of all parameters.  As such all plots can be viewed as incorporating the uncertainty from the detection model.  This is a lovely and natural way to present results and an advantage of having a joint model.  I'm not sure varprop can do this? [RJC - Varprop does do this through the posterior covariance $V_{\beta}$ matrix of the intercept, smoother and detection probability terms. But it doesn't do so in such an elegant way. Varprop is still reliant on a two-stage model.]

>>>>

% RJC 09/08/2019 Added case study text between % addtextstart and % addtextend markers.
% addtextstart
In this section, we apply the above methods to point transect data on the \hawaii{} \akepa{} (hereafter \akepa{}; \textit{Loxops coccineus}; nomenclature according to \citealp{usfws_akepa_1970}), an internationally and federally endangered Hawaiian honeycreeper (\citealp{usfws_akepa_1970, birdlife_akepa_2016}) that is endemic to \hawaii{} Island, USA.  Large-scale, quantitative surveys of Hawaiian forest birds and their habitat commenced in the mid-1970s through the Hawaii Forest Bird Survey (HFBS) \citep{scott_HFBS_1986}. Information from the HFBS was used to update the listing and delisting of endangered species, and establish preserves that coincided with native bird hotspots, including Hakalau Forest National Wildlife Refuge that was the first wildlife refuge with the primary purpose to protect, conserve and manage native forests for threatened and endangered bird and plant species. Due to the broad-scale coverage and robust design HFBS has become the baseline to determine changes in bird species distributions, population sizes and trends in density patterns over time.

During the 20th century \akepa{} declined dramatically due to habitat modification \citep{scott_HFBS_1986, pratt_avifaunal_1994},  mosquito-transmitted avian diseases \citep{pratt_avifaunal_1994, atkinson_wildlife_1995}, introduced predators \citep{lepson_akepa_1997}, and food resources competitors \citep{lepson_akepa_1997}. \Akepa{} has a global abundance of 16,248 (95\%CI 10,074\textendash25,198) birds that has been restricted to five spatially distinct populations \citep{judge_akepa_2018}. Hakalau Forest National Wildlife Refuge on \hawaii{} Island (hereafter Hakalau), supports the largest \akepa{} population that in 2012 was estimated at more than 11,000 birds \citep{camp_statespace_2016}. Maintaining and expanding the \akepa{} population at Hakalau is a primary conservation concern. Moreover, having unbiased and precise abundance estimates are required by land and resource managers for evaluating management actions and establishing management planning, and policy makers for decision-making processes.

\subsection*{Study area and sampling method}

[[Need to update this and the figure with the extended data area]]

Hakalau was established in 1985 to conserve 15,390-ha of montane forest habitat for native forest birds and rainforest plants. Annual forest bird surveys were initiated in 1987 to determine population status and track trends in abundance. Survey points were established along 15 transects following a systematic, random design with points approximately 150 m apart on transects located either 500 or 1,000 m apart. Following \cite{camp_population_2010, camp_statespace_2016}, we restrict our study area to the 3,061 ha open-forest stratum of Hakalau.  The open-forest stratum was previously heavily grazed, and since the removal of cattle in 1988 regeneration has proceeded naturally \citep{maxfield_hakalau_1998}. To the north the study area follows the refuge boundary while to the east it is bounded by a fence line (Fig. \ref{fig:2002studyareapointspt}). The southern boundary was modified from \cite{camp_population_2010} to exclude the non-sampled forested portion of their study area. The west side of the study area is bounded by pasture that is dominated by grass and is unsuitable habitat for \akepa.

@@

Surveys used point-transect methods, recording horizontal distances from survey points to individually detected birds. Surveys commenced at dawn and continued until 11:00 or halted when weather conditions exceeded prescribed conditions that hindered detecting birds (light rain, and wind and gust strength \textgreater Baufort scale 3). During 8-min counts trained observers recorded the species, exact distance to the nearest meter and detection type for each bird detected, along with the sampling conditions cloud cover, rain, wind strength, gust strength, and time of day each point was surveyed.   \cite{camp_population_2010,camp_statespace_2016} provide a detailed description of Hakalau, the open-forest study area and the bird surveys.

\subsection*{Data description}
[[ Need to update this with extended data ]]

For the purposes of our analysis we selected a single survey from the \akepa{} time series that contained on broad sampling of the study area with sufficient numbers of detections to estimate detectability. In 2002, 195 points were sampled using point-transect distance sampling methods within the 3,061 ha open-forest study area of Hakalau (Fig. \ref{fig:2002studyareapointspt}). On 67 points 152 \akepa{} were detected. The number of detections within each sampling unit ranged from zero to 6.

%\begin{figure}
%	\centering
%	\includegraphics[scale=0.5]{figures/2002counts_pt}
%	\caption{Counts of \akepa{} by point during the 2002 survey.}
%	\label{fig:2002countspt}
%\end{figure}

@@

In this section we introduce \texttt{inlabru} syntax and provide example code to demonstrate to specify the model.  First, we specify the SPDE model as
\begin{lstlisting}
matern <- inla.spde2.pcmatern(mesh, 
                              prior.sigma = c(2, 0.01),    
                              prior.range = c(130, 0.01))
\end{lstlisting}
which uses the penalised complexity priors for the hyperparameters of the GMRF as introduced by \cite{simpson_penalising_2017}, setting $\mathbb{P}(\sigma > 2) = 0.01$ and $\mathbb{P}(\rho < 130) = 0.01$ where $\rho$ is the range of the Mat\'ern covariance.

Second, we specify the thinning probability function.  The data contains only distances from the observer so we use equation \eqref{intensity-incomplete} to specify the log thinning probability function as
\begin{lstlisting}
thinning_function <- function(distance, lsig){
    log(2*pi) + log(distance) + log_halfnormal(distance, lsig)
    }
\end{lstlisting}
where \texttt{log\_halfnormal} is the log half-normal detection function $\log g(r | \sigma) = -1/2(r/\sigma)^2$ and the argument \texttt{lsig} is $\log(\sigma)$.  The dataset contains a column named \texttt{distance} that contains the observed distances.  The software will search for the appropriate columns using the names of the arguments specified in this function.  To allow the software to distinguish between names of covariates and parameters to be estimated we must also specify a model components object
\begin{lstlisting}
cmp <- ~ Intercept + grf(map = coordinates, model = matern) + lsig 
\end{lstlisting} 
This object names all model components to be estimated.  Note that the \texttt{grf} component is a name we have chosen and not a function.  The software allows user-specified names for the \texttt{f()} function known to users of \texttt{R-INLA}. Note that \texttt{coordinates} is a keyword that allows users to supply their data in the form of a spatial object using the \texttt{sp} package \citep{pebesma_spatial_2005}.

The model formula object is
\begin{lstlisting}
fml <- coordinates + distance ~ Intercept + grf +
	thinning_function(distance, lsig)
\end{lstlisting}
To fit the model we specify starting values for the parameters, the domain of possible observed distances and fit the model using the \texttt{lgcp} function
\begin{lstlisting}
starting_values <- data.frame(grf = 0, lsig = 3.65, Intercept = 0)
W <- 60   # transect radius
distance_domain <- seq(.Machine$double.eps, W, length.out = 30)
lgcp(components = cmp, 
     data = obs,
     samplers = samplers,
     domain = list(distance = distance_domain),
     formula = fml,
     options = list(result = starting_values, max.iter = 30))
\end{lstlisting}
where our observed data is contained in a \texttt{SpatialPointsDataFrame} object named \texttt{obs}.  Since not all sampling units observed birds we also specify the \texttt{samplers} object, also a \texttt{SpatialPointsDataFrame}, which contains the locations of all sampling units.

@@

The predicted mean of the posterior intensity is shown in \autoref{fig:intensity-mean-cv} along with a map of the coefficient of variation (CV) of the posterior intensity field on a regular prediction grid with cell area of approximately 1.7 hectares.  This shows a region of high intensity in the south and much lower intensity in the north.  The CV plot shows that the CV in the posterior intensity field is lower areas with greater sampling effort.  There is clear preferential sampling and in the north estimated intensities are lower which will also inflate the posterior CV.  It is difficult to know whether the high CV in the north is primarily caused by lower sampling effort or lower estimated abundance, or some combination of the two.

@@

The CV plot shows a measure of uncertainty relative to the mean intensity in each prediction grid cell.  However, often we are interested in the uncertainty around the mean intensity on the scale of the estimated intensity itself - perhaps seeking to identify areas where we can confidently predict particularly high or low intensities.  One approach attempt to communicate this uncertainty around the mean is to present the lower and upper quantiles corresponding to a 95\% credible interval for each prediction cell (\autoref{fig:intensity-quantiles}).

@@

This difference between summaries of the posterior and possible intensity surfaces that are consistent with the data becomes clear if one generates realizations of the posterior intensity field.  Three such realizations are in \autoref{fig:intensity-realizations}.  There are some notable differences when compared to the mean and quantile summary maps.  Each realization has finer-grained spatial structure than the mean or quantile plots of the posterior intensity.  In practice, this indicates that we should expect more spatial structure in our collected data than what we might expect if we only look at the posterior mean of the field.  The summary plots hide the spatial structure the model has been able to estimate from the data.  In general, the mean of a random field will contain less spatial structure than realisations of the field.

@@

The quantile plots are more concerning.  As mentioned above, the temptation is to interpret these as possible intensities that produced the observed data.  However, they in fact show the quantiles of the \textit{marginal} probability density for each prediction location. Presenting these quantiles side by side in a single map obscures the problem of multiple hypothesis testing.  To demonstrate this we treated the lower and upper quantile plots as though they were intensities and integrated them to obtain an expected abundance estimate of approximately 3,300 for the 0.025 quantile and 12,400 for the 0.975 quantile.  A naive (and tempting) interpretation of these numbers is as upper and lower limits of the 95\% credible interval for abundance. However, these abundance estimates are outside the support of the posterior for abundance (see \autoref{fig:realized-abundance-posterior}).  This emphasises that the temptation to treat these maps as showing possible intensity surfaces leads to interpretations (and possible management decisions) that are inconsistent with the very model that generated the maps.

As an alternative, we suggest using excursion sets and excursion functions \citep{bolin_excursion_2015}.  This is a recently developed approach to representing uncertainty in stochastic processes that specifically considers joint probability of events across multiple locations.  The positive excursion set with level $u$ for a function $f(s)$ with domain $\Omega$ is $A_u^{+}(f) = \{ s \in \Omega ; f(s) > u \}$ i.e. the set of all locations in $\Omega$ where $f$ exceeds the a threshold value $u$. For a stochastic process $\lambda(s)$ the positive excursion set with level $u$ and probability $1 - \alpha$ is
\begin{equation*}
E_{u,\alpha}^{+}(\lambda) = \argmax_{D}\{\lvert D \rvert : \mathbb{P}\left[D \subset A_u^{+}(\lambda)\right] \geq 1 - \alpha \}
\end{equation*}
It is important to note that $A_u^{+}(f)$ specifies a set for which a function $f(s)$ exceeds a threshold value $u$ for \textit{every location} in the set.  This means that when considering random quantities such as $\lambda(s)$ the definition of the positive excursion set $E_{u,\alpha}^{+}(\lambda)$ is the largest such set for which a threshold is exceeded simultaneously for all locations in the set.  Because $\lambda(s)$ is stochastic we can only ever say this is true with some probability and use something akin to a confidence level by specifying $\alpha$.  Excursion sets can be estimated by considering candidate sets for $D$ of increasing size and a sequential integration scheme to estimate probabilities.  An implementation is available in the \texttt{excursions} package \citep{bolin_calculating_2018} available through the Comprehensive R Archive Network \citep{r_2017}.

\autoref{fig:excursions} (left panel) shows the positive excursion set with a level corresponding to 1 bird per hectare with probability 0.95.  This figure can be interpreted in a natural way.  It is the largest possible region for which the intensity is greater than 1 bird per hectare for every location within the region, with probability 0.95.
\begin{figure}
	\includegraphics[scale=0.5]{figures/excursions.png}
	\caption{Left:  The positive excursion set with a level corresponding to 1 bird per hectare and probability 0.95.  Right: The positive excursion function with a level corresponding to 1 bird per hectare}
	\label{fig:excursions}
\end{figure}
To visualise multiple such maps we can use the excursion function $F_u^{+}(s) = \sup \{1 - \alpha ; s \in E_{u,\alpha}^+ \}$, which defines for each location something similar to a p-value.  For each $s$ it is the largest possible probability $1 -\alpha$ for which that location would be in the excursion set defined using probability $1 - \alpha$.  The excursion function with a level corresponding to 1 bird per hectare is shown in \autoref{fig:excursions} (right panel).

@@

A key feature of these plots is that they display either discrete sub-regions of the area of interest (for excursion sets) or the probability of membership of such discrete sub-regions (for excursion functions).  We believe the risk of interpreting these maps as intensities is much lower than for the quantile plots.  The fact that it is possible to estimate discrete regions may also be useful to communicate results to conservation practitioners, land and resource managers and policy decision-makers.  We anticipate the excursion function will be most useful in communicating to audiences with statistical training. % [RJC - I think statistically savvy audiences will readily grasp the excursion functions. Do you mean "without" statistical training?]

Another key aim of performing distance sampling surveys is to estimate the abundance of the population of interest.  We describe here how to estimate the posterior for the abundance. Let $n$ denote the abundance of the population within the region of interest and $N$ its corresponding random variable within our Bayesian framework.  Integrating the mean of the random field provides a point estimate for the abundance, which we call the \textit{expected} abundance.  Here we demonstrate the importance of considering the full posterior for $N$ which we call the \textit{realised} abundance.

Let $N = \int_{\Omega}\lambda(s)\mathrm{d}s$. Note that $N$ is a random variable and the integral of $\mathbb{E}[\lambda(s)]$ is equivalent to $\mathbb{E}(N)$ by Fubini's theorem, hence the term expected abundance.  As we mentioned above, this is the most sensible point estimate for the abundance.  To communicate uncertainty around the expected abundance we estimate the posterior $\pi(N | \lambda)$ by a monte-carlo method.  Taking $m$ monte-carlo samples  $\lambda_1, \ldots, \lambda_m$ of the posterior intensity field we estimate the posterior for the abundance as $\pi(N | \lambda) \approx 1 / m \sum_{i=1}^m \pi (N | \lambda = \lambda_i)$. This is well-defined since $\pi(N | \lambda = \lambda_i)$ is a Poisson probability mass function with rate parameter $\int_{\Omega}\lambda_i(s)\mathrm{d}s$. \autoref{fig:realized-abundance-posterior} shows the approximate posterior for $N$ with $m = 20,000$.  This allows us to estimate the probability for any specific value of realised abundance $n$ by calculating $\pi(N = n | \lambda)$.  For reference, \autoref{fig:realized-abundance-comparison} shows a single Poisson PMF with rate parameter equal to the expected abundance.

\autoref{fig:half-normal} shows a summary of the posterior for the detection function, plotting the posterior mean along with 95\% credible intervals.  We note again that all model summaries shown here are based on monte-carlo sampling of the joint posterior of all model parameters.  As such all the plots incorporate the uncertainty in the detection process.

@@

\begin{figure}
	\includegraphics[scale=0.6]{figures/halfnormal.png}
	\caption{Posterior detection function}
	\label{fig:half-normal}
\end{figure}

@@
